### 2014-04-19 patchy-selection-paper.tex
.

We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
as the sum of independent migrant families, each of whose dynamics are given by a spatial branching process.
Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\delta B(r_0)$ its boundary.
Suppose that there is a constant outflux of migrant families of rate $\gamma(x)$ from each point $x \in \delta B(r_0)$,
e.g.\ with locations and times of appearances as the points of a Poisson process.
Roughly speaking, equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}
differ only by substituting the average number of members of a migrant family in the new patch
for the probability that a family has any members in the new patch;
by the discussion of the previous section, these differ only by a constant factor.

First we will make this idea more formal.
Suppose that $Z$ is one such spatial branching process as above, started with a single individual at $x$,
defined so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the set $A$.
The \emph{mean occupation density} of $Z$ is then defined to be the density of $\E[\int_0^t Z_t dt]$,
i.e.\ the smooth function $u(x,y)$ that, integrated over a set $A$, 
gives the expected total number of individuals ever living in $A$:
\begin{align} \label{eqn:occupation_density_defn}
    \E\left[ \int_0^\infty Z_t(A) dt \right] = \int_A u(x,y) dy .
\end{align}
We can think of $u(x,y)$ as the expected total number of offspring of a family beginning at $x$
that ever live at $y$.
By independence of reproduction and spatial motion,
$u(x,y)$ is the sum over all times of the expected number of offspring
multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
i.e.\
\begin{align}
    u(x,y) = \int_0^t e^{-\alpha t} p_t(x,y) dt .
\end{align}
Now, \eqref{eqn:gestalt_q} is
\begin{align}
    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  \\
         &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-\alpha t} p_t(x,y) dt dx .
\end{align}

Now consider the probability that a migrant family from $x$ will ever establish in the new patch,
which we assume has relatively small area $A$ and is located near position $y$.
The probability that this family reaches $A$ by time $t$ is $(1-p_e(t))$
multiplied by the probability that the family reaches the patch, given survival.
If $t$ is large then the previous section argues that $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$,
and the probability of reaching the patch by $t$ approximately the chance that a Brownian path
will have hit the patch by then.
If $w = A^{1/d}$ is the width of the patch, then

>>>>

We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
as the sum of independent migrant families, each of whose dynamics are given by a spatial branching process.
Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
Suppose that there is a constant outflux of migrant families of rate $\gamma(x)$ from each point $x \in \partial B(r_0)$.
To make expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} more concrete,
we need to replace ``outflux of families'' by a sum over points at distance $r_0$ from the patch
of the outflux $\gamma$ at that point, mulitplied by the relevant second quantity.
By examining the constituent terms,
it will turn out that \eqref{eqn:gestalt_migrate} is a constant multiple of \eqref{eqn:gestalt_q}.

First consider \eqref{eqn:gestalt_q}.
Suppose that $Z$ is one such spatial branching process as above, started with a single individual at $x$,
defined so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the set $A$.
The \emph{mean occupation density} of $Z$ can be thought of 
as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
Concretely, it is the density of $\E[\int_0^t Z_t dt]$,
i.e.\ the smooth function $u(x,y)$ that, integrated over a set $A$, 
gives the expected total number of individuals ever living in $A$:
\begin{align} \label{eqn:occupation_density_defn}
    \E\left[ \int_0^\infty Z_t(A) dt \right] = \int_A u(x,y) dy .
\end{align}
Since reproduction and spatial motion are independent,
$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
i.e.\
\begin{align}
    u(x,y) = \int_0^t e^{-\alpha t} p_t(x,y) dt .
\end{align}
The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
is the sum of mean occupation densities of a constant outflux of branching processes
from $\partial B(r_0)$:
\begin{align} \label{eqn:occupation_integral}
    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  % \\
         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-\alpha t} p_t(x,y) dt dx .
\end{align}

Now we turn to the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch,
which we assume is circular, has relatively small area $A$ and is located near position $y$.
% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-p_e(t))$,
% multiplied by the probability that the family reaches the patch, given survival.
% If $t$ is large then as in the previous section, $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$.
It would be possible to analyze directly the probability that a given migrant family establishes in $A$,
as in \citet{barton1987probability,branchingbrownian};
but we make a simpler approximation, approximating this by the chance that the ``trunk'' hits the new patch,
multiplied by the chance that at least member of the family escapes demographic stochasticity to establish.
Write $h(x,y)$ for the probability that the Brownian trunk
hits the new patch centered at $y$ before the family dies out,
and $f(A)$ for the chance that the family manages to establish in the new patch,
given that it successfully arrives.
As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is an integral of $h(x,y)$:
\begin{align}\label{eqn:migrate_integral}
  \migrate(y) &= \int_{\partial B(r_0)} \gamma(x) h(x,y) f(A) dx .
\end{align}

Intuitively, $h(x,y)$ counts each family once if it hits the patch,
while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
Since the expected total time spent in the patch does not depend on the distance at which the family started
if we know that it survived to hit the patch,
it is intuitively reasonable that $u(x,y)$ and $h(x,y)$ are simply related.
Using this fact,
and that $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$,
it turns out (see Appendix~\ref{apx:hitting_occupation}) that
\begin{align} 
  h(x,y) = \frac{ s(A) }{ \E[K] } u(x,y),
\end{align}
where $s(A)$ is the total expected occupation time of the patch given the family arrives,
and recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a constant factor
(at least asymptotically for large distances),
we now know that the outflux of successful migrant families is equal to the mean occupation frequency
multiplied by $s(A) f(A) / \E[K]$.
The probability $f(A)$ that the family does not die out once it arrives at the new patch
is approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$.
% If we let $w$ be the width of the patch, then the remaining term is
% \begin{align}
%   s(w) \simeq \frac{1}{2\alpha} \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
% \end{align}
Combining these terms with the expression for $s(A)$ from Appendix~\ref{apx:hitting_occupation},
\begin{align}
  \migrate(y) \approx q(y) \times \frac{ s_p  }{ \alpha \xi^2 } 
  \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
\end{align}
